---
title: "Supplement to: Multiple imputation of missing covariates when using the Fine--Gray model"
author: 
  - name: E. F. Bonneville
    affil-id: 1
  - name: J. Beyersmann
    affil-id: 2
  - name: R. H. Keogh
    affil-id: 3
  - name: J. W. Bartlett
    affil-id: 3
  - name: T. P. Morris
    affil-id: 4
  - name: N. Polverelli
    affil-id: 5
  - name: L. C. de Wreede
    affil-id: 1,6,*
  - name: and H. Putter
    affil-id: 1,7,*
affiliations:
  - id: 1
    name: Department of Biomedical Data Sciences, Leiden University Medical Center, the Netherlands
  - id: 2
    name: Institute of Statistics, Ulm University, Germany
  - id: 3
    name: Department of Medical Statistics, London School of Hygiene and Tropical Medicine, United Kingdom
  - id: 4
    name: MRC Clinical Trials Unit at UCL, United Kingdom
  - id: 5
    name: Unit of Bone Marrow Transplantation, Division of Hematology, Fondazione IRCCS Policlinico San Matteo di Pavia, Italy
  - id: 6
    name: DKMS Clinical Trials Unit, Germany
  - id: 7
    name: Mathematical Institute, Leiden University, the Netherlands
  - id: "*"
    name: Shared senior authorship
format: 
  pdf:
    documentclass: article
    number-sections: true
    code-block-border-left: false
    papersize: a4
    geometry:
      - margin=1in
    template-partials:
      - title.tex
    include-in-header:
      text: |
        \usepackage[noblocks]{authblk}
        \renewcommand*{\Authsep}{, }
        \renewcommand*{\Authand}{, }
        \renewcommand*{\Authands}{, }
        \renewcommand\Affilfont{\small}
        \usepackage{booktabs}
        \usepackage{array}
        \usepackage{multirow}
        \usepackage{longtable}
        \renewcommand{\thesection}{S\arabic{section}}
execute:
  echo: true
editor_options: 
  chunk_output_type: console
---


```{r}
#| label: setup
#| warning: false
#| echo: false
#| fig-cap-location: top

# Add this to quarto pipeline after
source(here::here("packages.R"))
invisible(lapply(list.files(here::here("R"), full.names = TRUE), source))
options(contrasts = rep("contr.treatment", 2))

# Theme tings
cols <- c("#CABEE9", "#7C7189", "#FAE093", "#D04E59", "#BC8E7D", "#2F3D70")
theme_set( # set base size to 14 instead of 16
  theme_light(base_size = 16, base_family = "Roboto Condensed") +
    theme(
      strip.background = element_rect(fill = cols[2], colour = "white"),
      strip.text = element_text(colour = 'white')
    )
)
```

# Minimal code example

This is the minimal `R` code companion to section 3.4 of main manuscript. The parameters from the simulation study scenario with $p = 0.15$, random censoring, and correctly specified Fine--Gray were used to generate the example dataset below.

```{r}
#| echo: false

# Generate dataset
set.seed(202405)

params <- list(
  "cause1" = list(
    "formula" = ~ X + Z,
    "betas" = c(0.75, 0.5),
    "p" = 0.15,
    "base_rate" = 1,
    "base_shape" = 0.75
  ),
  "cause2" = list(
    "formula" = ~ X + Z,
    "betas" = c(0.75, 0.5),
    "base_rate" = 1,
    "base_shape" = 0.75
  )
)

dat_minimal <- generate_dataset(
  n = 2000,
  args_event_times = list(
    mechanism = "correct_FG",
    params = params,
    censoring_type = "exponential"
  ),
  args_missingness = list(
    mech_params = list(
      "prob_missing" = 0.4,
      "mechanism_expr" = "1.5 * Z"
    )
  )
)
dat_minimal[, ':=' (time = round(time, digits = 6), Z = round(Z, digits = 3))]
dat <- data.frame(dat_minimal[order(id), c("id", "time", "D", "X", "Z")])
rownames(dat) <- NULL

#See suppl of https://onlinelibrary.wiley.com/doi/full/10.1002/sim.8387
```


```{r loading}
#| warning: false

# Load libraries
library(data.table)
library(survival)
library(kmi)
library(mice)
library(smcfcs)

# Minimal dataset
head(dat, n = 10)
sapply(dat, class)
nrow(dat)
```

1. Add columns $\hat{H}_1(T)$ and $\hat{H}_2(T)$ to the original data, which are the marginal cause-specific cumulative hazards for each competing risk evaluated at an individual's event or censoring time (obtained using the Nelson--Aalen estimator).

```{r}
# Add cause-specific event indicators + cumulative hazards
dat$D1 <- as.numeric(dat$D == 1)
dat$D2 <- as.numeric(dat$D == 2)
dat$H1 <- mice::nelsonaalen(data = dat, timevar = "time", statusvar = "D1")
dat$H2 <- mice::nelsonaalen(data = dat, timevar = "time", statusvar = "D2")
```

2. Multiply impute the potential censoring for those failing from cause 2 using \{kmi\}, yielding $m$ censoring complete datasets (i.e. with "complete" $V$). Any completely observed covariates that are known to affect the probability of being censored should be included as predictors in the model for the censoring process. \{kmi\} imputes based on stratified Kaplan--Meier when $Z$ are categorical, and based on a Cox model when at least one of $Z$ are continuous.

<!-- Mention administrative censoring here -->

```{r}
# Multiply impute the censoring times
cens_imps <- kmi(
  formula = Surv(time, D != 0) ~ 1, # Additional predictors added here
  data = dat,
  etype = D,
  failcode = 1, # Specify event of interest
  nimp = 5
)
```

3. In each censoring complete dataset, add an additional column $\hat{\Lambda}_1(V)$. This takes the value of the marginal cumulative subdistribution hazard for cause 1 at an individual's observed or imputed subdistribution time, obtained with the Nelson--Aalen estimator based on $I(D = 1)$ and imputed $V$.

```{r}
# Preparation for covariate imputation:
list_to_impute <- lapply(cens_imps$imputed.data, function(imp_dat) {
  
  # Adjust to new ordering from kmi (cause 2 individuals appended at bottom)
  dat_to_impute <- cbind(cens_imps$original.data, imp_dat)
  
  # Compute/add Lambda_1(V) in each imputed dataset
  dat_to_impute$Lambda1 <- mice::nelsonaalen(
    data = dat_to_impute, 
    timevar = "newtimes", # kmi naming for V
    statusvar = "D1" # I(D=1)
  )
  return(dat_to_impute)
})

# newevent is equal to I(D=1)
head(list_to_impute[[1]])
```

4. In each censoring complete dataset (each with different $V$ and $\hat{\Lambda}_1(V)$, but same $\hat{H}_1(T)$ and $\hat{H}_2(T)$), create a single imputed dataset using the desired covariate imputation method(s).

```{r}
# Prepare predictor matrices for MICE using the first censoring complete dataset
predmat_cs_approx <- predmat_fg_approx <- mice::make.predictorMatrix(list_to_impute[[1]])
predmat_cs_approx[] <- predmat_fg_approx[] <- 0
predmat_cs_approx["X", c("Z", "D1", "D2", "H1", "H2")] <- 1
predmat_fg_approx["X", c("Z", "D1", "Lambda1")] <- 1
predmat_fg_approx

# Prepare the methods:
# - Approx methods: model type for X | Z, outcome
methods_approx <- mice::make.method(list_to_impute[[1]])
# - SMC methods: proposal model for X | Z (need to use {smcfcs} naming)
methods_smcfcs <- mice::make.method(
  list_to_impute[[1]],
  defaultMethod = c("norm", "logreg", "mlogit", "podds")
)
methods_smcfcs
```

```{r}
#| label: step_4
#| results: false

# Impute X
list_imps <- lapply(list_to_impute, function(imp_dat) {

  m <- 1
  iters <- 10
  
  imps_cs_approx <- mice(
    data = imp_dat,
    m = m,
    maxit = iters,
    method = methods_approx,
    predictorMatrix = predmat_cs_approx
  )

  imps_fg_approx <- mice(
    data = imp_dat,
    m = m,
    maxit = iters,
    method = methods_approx,
    predictorMatrix = predmat_fg_approx
  )

  imps_cs_smc <- smcfcs(
    originaldata = imp_dat,
    smtype = "compet",
    smformula = list(
      "Surv(time, D == 1) ~ X + Z",
      "Surv(time, D == 2) ~ X + Z"
    ),
    method = methods_smcfcs,
    m = m,
    numit = iters
  )

  imps_fg_smc <- smcfcs(
    originaldata = imp_dat,
    smtype = "coxph",
    smformula = "Surv(newtimes, D1) ~ X + Z",
    method = methods_smcfcs,
    m = m,
    numit = iters
  )

  # Bring all the imputed datasets together
  imps <- rbind.data.frame(
    cbind(method = "CCA", imp_dat),
    cbind(method = "cs_smc", imps_cs_smc$impDatasets[[1]]),
    cbind(method = "cs_approx", complete(imps_cs_approx, action = 1L)),
    cbind(method = "fg_smc", imps_fg_smc$impDatasets[[1]]),
    cbind(method = "fg_approx", complete(imps_cs_approx, action = 1L))
  )
  return(imps)
})
```

<!-- Note about assessing convergence -->

5. Fit the Fine--Gray substantive model in each imputed dataset (using standard Cox software with $I(D = 1)$ and imputed $V$ as outcome variables), and pool the estimates using Rubin's rules.

```{r}
#| label: step_5

# Bind everything together
dat_imps <- rbindlist(list_imps, idcol = ".imp")
dat_imps

# To use the usual workflow: subset one of the methods first
imps_fg_smc <- dat_imps[method == "fg_smc"]

# Fit model in each imputed dataset
mods_fg_smc <- lapply(
  X = split(x = imps_fg_smc, f = imps_fg_smc$.imp), 
  FUN = function(imp_dat) coxph(Surv(newtimes, D1) ~ X + Z, data = imp_dat)
)

# Pool results
summary(pool(mods_fg_smc))

# Otherwise: use (nested) data.table workflow to pool for all methods simultaneously
dat_mods <- dat_imps[, .(
  mod = list(coxph(Surv(newtimes, D1) ~ X + Z, data = .SD))
), by = c("method", ".imp")]
dat_mods

dat_mods[, summary(pool(as.list(mod))), by = "method"]
```

\newpage

# Applied data example

```{r}
#| echo: false
#| results: false

# Load in the good stuff
tar_load(applied_dat, store = here::here("_targets"))
tar_load(applied_dat_pooled, store = here::here("_targets"))
dat <- applied_dat$dat
sm_predictors <- applied_dat$sm_predictors

#
preds_final <- applied_dat_pooled$pooled_cumincs
method_labs <- c(
  "Compl. cases" = "CCA",
  "SMC-FCS cause-spec" = "CS-SMC",
  "MICE cause-spec" = "CS-Approx",
  "SMC-FCS Fine-Gray" = "FG-SMC",
  "MICE subdist" = "FG-Approx"
)
preds_final[, method := factor(
  method, levels = names(method_labs), labels = method_labs
)]

# First make the cuminc figure
p_cuminc <- preds_final |>
  ggplot(aes(times, p_pooled, group = method)) +
  #geom_ribbon(aes(ymin = CI_low, ymax = CI_upp, fill = method), alpha = 0.5) +
  geom_step(aes(col = method, linetype = method), linewidth = 1) +
  scale_color_manual(values = cols[c(1, 2, 6, 4, 5)]) +
  coord_cartesian(ylim = c(0, 0.225)) +
  labs(
    col = "Method",
    linetype = "Method",
    x = "Time since alloHCT (months)",
    y = "(Pooled) Baseline cumulative incidence"
  )

p_ci_width <- preds_final |>
  ggplot(aes(times, CI_width, group = method)) +
  geom_step(aes(col = method, linetype = method), linewidth = 1) +
  scale_color_manual(values = cols[c(1, 2, 6, 4, 5)]) +
  coord_cartesian(ylim = c(0, 0.225)) +
  labs(
    col = "Method",
    linetype = "Method",
    x = "Time since alloHCT (months)",
    y = "(Pooled) 95% Confidence interval width"
  )


p_comb <- p_cuminc + p_ci_width & xlab(NULL) & theme(legend.position = "top")

p_final <- wrap_elements(panel = p_comb + plot_layout(guides = "collect")) +
  labs(tag = "Time since alloHCT (months)") +
  theme(
    plot.tag = element_text(size = rel(1)),
    plot.tag.position = "bottom"
  )

ggplot2::ggsave(
  plot = p_final,
  filename = here::here("analysis/figures/applied_base_cuminc.pdf"),
  width = 11,
  scale = 1,
  height = 7,
  device = cairo_pdf
)
```

## Data dictionary

```{r}
#| echo: false
#| warning: false

library(gtsummary)
library(kableExtra)

# Tbl summary
df <- applied_dat$dat
df_predz <- copy(df[, ..sm_predictors]) #|>
# Re-transform for descriptives
df_predz[, ':=' (
  age_allo1_decades = (age_allo1_decades + 6) * 10,
  year_allo1_decades = year_allo1_decades * 10 + 10 + 2009,
  intdiagallo_decades = intdiagallo_decades * 10,
  pb_allo1 = pb_allo1 * 5,
  hb_allo1 = hb_allo1 * 5 + 10,
  wbc_allo1 = exp(wbc_allo1 + log(15.1))
)]

# Edit some factor labels
levels(df_predz$cmv_match)[1] <- "Patient negative/Donor negative"
levels(df_predz$donrel_bin)[1] <- "HLA identical sibling"
levels(df_predz$hctci_risk) <- c(
  "Low risk ($0$)", "Intermediate risk ($1-2$)", "High risk ($\\geq 3$)"
)
levels(df_predz$KARNOFSK_threecat) <- c("$\\geq 90$", "$80$", "$\\leq 70$")
levels(df_predz$ric_allo1) <- c("Standard", "Reduced")
levels(df_predz$ruxo_preallo1) <- levels(df_predz$tceldepl_bin) <- c("No", "Yes")
levels(df_predz$submps_allo1) <- c("Primary MF", "Secondary MF")

# Edit some variable labels
attr(df_predz$hctci_risk, "label") <- "HCT-CI risk category"
attr(df_predz$KARNOFSK_threecat, "label") <- "Karnosfky performance score"
attr(df_predz$tceldepl_bin, "label") <- "T-cell depletion (in- or ev-vivo)"
attr(df_predz$ruxo_preallo1, "label") <- "Ruxolitinib given"
attr(df_predz$age_allo1_decades, "label") <- "Patient age (years)"
attr(df_predz$submps_allo1, "label") <- "Disease subclassification"
attr(df_predz$cmv_match, "label") <- "Patient/donor CMV match"
attr(df_predz$intdiagallo_decades, "label") <- "Interval diagnosis-transplantation (years)"
attr(df_predz$year_allo1_decades, "label") <- "Year of transplantation"
attr(df_predz$pb_allo1, "label") <- "Peripheral blood (PB) blasts (%)"
attr(df_predz$wbc_allo1, "label") <- "White blood cell count (WBC, x$10^9$/L)"
attr(df_predz$WEIGLOSS_allo1, "label") <- ">10% Weight loss prior to transplantation"

data_dict_caption <- "Data dictionary. CMV: cytomegalovirus; HLA: human leukocyte antigen; HCT-CI: Hematopoietic stem cell transplantation-comorbidity index; MF: myelofibrosis."

tbl_test <- tbl_summary(
  df_predz,
  missing_text = "(Missing)",
  type = all_dichotomous() ~ "categorical"
) |>
  as_kable_extra(
    format = "latex",
    booktabs = TRUE,
    longtable = TRUE,
    caption = data_dict_caption,
    linesep = "",
    addtl_fmt = FALSE
  )

gsub(tbl_test, pattern = "%", replacement = "\\\\%")
```

## Non-parametric cumulative incidence curves

```{r}
#| fig-width: 9
#| fig-height: 7
#| dev: cairo_pdf
#| echo: false
#| fig-pos: 'H'
#| fig-cap: "Stacked non-parametric cumulative incidence curves for competing relapse and non-relapse mortality, in dataset of 3982 primary and secondary myelofibrosis patients."
#| fig-cap-location: top



np_curves <- prodlim(Hist(time_ci_adm, status_ci_adm) ~ 1, data = dat)

par(family = "Roboto Condensed")
plot(
  np_curves,
  cause = "stacked",
  col = cols[c(4, 6)],
  ylab = "Stacked cumulative incidence",
  lty = c(1, 2),
  xlab = "Time since alloHCT (months)",
  percent = FALSE,
  atrisk.at = seq(0, 60, by = 10),
  atrisk.col = "black",
  atrisk.title = "At risk:",
  legend = FALSE
)
legend(
  x = 0, y = 0.95,
  legend = c("Relapse", "Non-relapse mortality"),
  lty = c(1, 2),
  col = cols[c(4, 6)],
  bty = 'n',
  lwd = c(3, 3)
)
```

## Pooled regression coefficeints

```{r}
#| echo: false

pooled_mods <- applied_dat_pooled$pooled_coefs
pooled_mods[, summ := paste0(
  round(estimate, 2), " [", round(conf.low, 2), ", ", round(conf.high, 2), "]"
)]
pooled_mods[, method := factor(method, levels = names(method_labs), labels = method_labs)]

df_tbl <- dcast(pooled_mods, term + method ~ mod_type, value.var = "summ")
df_tbl[, term := factor(
  term,
  levels = c(
    "ric_allo1reduced",
    "cmv_matchOther",
    "vchromos_preallo1Abnormal",
    "donrel_binOther",
    "hb_allo1",
    "hctci_riskintermediate risk (1-2)",
    "hctci_riskhigh risk (>= 3)",
    "intdiagallo_decades",
    "KARNOFSK_threecat80",
    "KARNOFSK_threecat<80",
    "submps_allo1sMF",
    "sweat_allo1Yes",
    "age_allo1_decades",
    "PATSEXMale",
    "pb_allo1",
    "ruxo_preallo1yes",
    "tceldepl_binyes",
    "wbc_allo1",
    "WEIGLOSS_allo1Yes",
    "year_allo1_decades"
  ),
  labels = c(
    "Conditioning: reduced",
    "CMV match: other",
    "Cytogenetics: abnormal",
    "Donor relation: other",
    "Hemoglobin (per $5$ g/dL)",
    "HCT-CI ($1-2$)",
    "HCT-CI ($\\geq 3$)",
    "Interval diagnosis to alloHCT (decades)",
    "Karnofsky ($80$)",
    "Karnofsky ($\\leq 70$)",
    "Disease subclassification: secondary MF",
    "Night sweats: yes",
    "Patient age (decades)",
    "Patient sex: male",
    "PB Blasts (per $5$\\%)",
    "Ruxolitinib given: yes",
    "T-cell depletion: yes",
    "WBC count (log)",
    "Weight loss: yes",
    "Year of alloHCT (decades)"
  )
)]

tb_breaks <- rep(5, length(levels(df_tbl$term)))
names(tb_breaks) <- levels(df_tbl$term)

pooled_coefs_caption <- "Pooled log hazard ratios [log HR, 95\\% confidence interval] for Fine--Gray model for relapse, cause-specific Cox model relapse, and cause-specific Cox model for non-relapse mortality (NRM)."

df_tbl[order(term)][, !c("term")] |>
  kbl(
    format = "latex",
    longtable = TRUE,
    booktabs = TRUE,
    caption = pooled_coefs_caption,
    align = c("l", "r", "r", "r"),
    col.names = c(
      "Term + method",
      "Relapse subdist.~log HR",
      "Relapse cause-spec.~log HR",
      "NRM cause-spec.~log HR"
    ),
    escape = FALSE
  ) |>
  # Important that pack_rows() before the styling
  pack_rows(index = tb_breaks, escape = FALSE) |> 
  kable_styling(
    latex_options = c("repeat_header"),
    repeat_header_continued = TRUE,
    repeat_header_method = "replace",
    font_size = 9
  )
```
